Binocular integration of chromatic and luminance signals

### 2024-04-01 manuscript.tex
Trying again

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Binocular integration of chromatic and luminance signals}
\author{Daniel H. Baker, Kirralise J. Hansford, Federico G. Segala, Annie Y. Morsi,\\
Rowan J. Huxley, Joel T. Martin, Maya Rockman \& Alex R. Wade}
\date{Department of Psychology, University of York}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

@@

A useful framework for understanding binocular signal processing is the two-stage gain control model of binocular combination introduced by Meese et al. (2006). This model features interocular suppression between monocular channels, followed by binocular summation. The model accounts well for the pattern of contrast discrimination (`dipper') functions for four distinct ocular configurations (see also Georgeson et al., 2016), illustrated in Figure \ref{fig:exampledips}. In the monocular condition, participants must discriminate between stimuli of two contrasts (a `pedestal', and a `pedestal plus target') that are both presented to one eye, whilst the other eye views mean luminance. Threshold is defined as the minimum target contrast required to make this judgement with 75\% accuracy. The binocular condition is the same, except that the same stimuli are shown to both eyes. In the half-binocular condition the pedestal is shown to both eyes, but the target increment shown only to one eye. Finally, the dichoptic condition involves presenting the pedestal to one eye, and the target increment to the other eye.

@@

{\centering \includegraphics{Figures/exampledips}

@@

\caption{Illustration of stimulus conditions (left) and example dipper functions (right).}\label{fig:exampledips}
\end{figure}

@@

In Experiments 1 and 2, the stimuli were horizontal sinusoidal gratings with a spatial frequency of 1c/deg (see examples in Figure \ref{fig:examplestim}). The gratings were windowed by a raised cosine envelope with a diameter of 3 degrees. Spatial phase, relative to a central fixation cross, was randomised on each trial across the four cardinal phases. In the achromatic conditions, the sine-wave modulated all three monitor colour channels equally (red, green and blue). In the L-M condition, we generated isoluminant stimuli for each participant (see Procedures) designed to maximise contrast between L and M cones, whilst keeping S cone activity constant. In the S-(L+M) condition, the isoluminant stimuli maximised S-cone contrast. Stimuli were converted from cone space to monitor RGB coordinates using the monitor spectral readings and the Stockman-Sharpe 2 degree cone fundamentals (Stockman \& Sharpe, 2000). The stimuli in Experiment 3 were temporal modulations of luminance applied to a disc made using the same raised cosine envelope as described above, but with no further spatial modulation. The stimuli counterphase flickered sinusoidally at 4Hz (see examples in the lower portion of Figure \ref{fig:examplestim}). In all experiments, we displayed a binocular fusion lock, consisting of three concentric rings of small square elements with random colour. A black central fixation cross was also displayed throughout.

@@

{\centering \includegraphics[width=0.5\linewidth]{Figures/examplestim}

@@

\caption{Example stimuli. Upper row shows grating stimuli used in Experiments 1 and 2. Lower row shows one cycle of sinusoidal flicker applied to a uniform disc. Note that the rendering of all stimuli will depend on the device used to display or print this image, and so the chromatic stimuli are unlikely to appear isoluminant, and there may be additional luminance nonlinearities that were not present in the stimuli displayed during the experiments.}\label{fig:examplestim}
\end{figure}

@@

Before beginning primary data collection, each participant in Experiments 1 and 2 completed an isoluminance adjustment task. Grating stimuli were presented that counterphase flickered at 5Hz, defined about either the L-M or S-(L+M) plane in cone space. Participants used a trackball to dynamically adjust the colour angle of the stimulus to minimise the percept of flicker. Each participant completed ten such trials for each colour plane, and the average angle across repetition was taken as the isoluminant point, and used to generate stimuli for the main experiment for that participant. Settings were very similar across participants for the S-(L+M) direction, and somewhat more heterogeneous for the L-M direction (see Figure \ref{fig:isofig}).

@@

{\centering \includegraphics{Figures/isosettings}

@@

\caption{Isoluminance settings from all participants in Experiments 1 and 2. Panel (a) shows red/green and panel (b) shows blue/yellow settings that were subsequently used to generate stimuli in the main experiments. Within each panel, solid lines show the mean settings for each participant, and black curves show the range of possible stimuli displayed during the adjustment task.}\label{fig:isofig}
\end{figure}

In Experiment 1, participants completed a two-interval-forced-choice (2IFC) contrast discrimination task. Stimuli were presented for 200ms, with an interstimulus interval of 400ms. Each interval was indicated by an auditory beep, and participants made their responses using a two-button trackball. Correct responses were followed by a high pitched tone, and incorrect responses by a low pitched tone. Each block of the experiment tested a single pedestal contrast level, and lasted around 12 minutes. On each trial the target contrast level was determined by a 3-down-1-up staircase procedure. There were 8 interleaved staircases in total; four stimulus arrangements (see Figure \ref{fig:exampledips}) combined factorially with two target eye assignments. Each pedestal contrast was repeated 3 times by each participant, and the block order was randomised. The experiment lasted around 4 hours per participant for each chromatic condition, and took place over the course of several weeks. In total, the experiment consisted of 67978 trials (pooled across participants).

@@

We also implemented a Bayesian hierarchical version of the model using the Stan language (Carpenter et al., 2017). This used a binomial generator function to model the proportion correct data at each target level, and was fit simultaneously to all participants for a given experiment, but separately for each chromatic condition of Experiment 1, and the flickering disc data from Experiment 3 (i.e.~four fits in total, as for the simplex fitting). Prior distributions were Gaussian, with means determined from published values (see first row of Table \ref{tab:parametertable}). This modelling primarily focuses on examining posterior parameter distributions, rather than a model comparison approach. We generated over 1 million posterior samples for the model, and retained 10\% of them for plotting.

@@

where \emph{AC} represents the achromatic contrast, \emph{RG} represents the red/green contrast, \emph{BY} represents the blue/yellow contrast, and \(\omega_A\), \(\omega_R\) and \(\omega_B\) are the accompanying weights of interocular suppression. There is an equivalent expression for the right eye, and for each of the two isoluminant chromatic pathways. To simplify the model, and avoid free parameters that are poorly constrained by the data, we fixed several parameters at the average values from the fits from Experiment 1 (Table \ref{tab:parametertable}), such that \(p=7.82\), \(q=6.4\), \(m=1.21\), \(S=1.03\) and \(k=0.23\). We additionally fixed the within-mechanism weight of interocular suppression (\(\omega\)) at 1. This left 9 free parameters: a \(Z\) parameter for each mechanism, and six cross-mechanism weights of interocular suppression. These parameters were again estimated within a Bayesian hierarchical framework, using the trial-wise data from Experiment 2. Note that the model as specified does not currently include monocular suppression between different pathways, as we did not collect any data for these conditions. Previous work (Chen et al., 2000) has measured such interactions, and they could in principle be incorporated into the denominator of either stage 1 or stage 2 in the model.

@@

All experimental code, raw data and analysis scripts are available at: \url{https://osf.io/3vdga/}. The linked GitHub repository also contains a fully reproducible version of the manuscript. Note that we deviated slightly from the planned preregistration, in that we did not collect data for chromatic flickering discs, or for the cross-pathway dichoptic experiment using disc stimuli. This is because the grating data from Experiments 1 and 2, and the achromatic disc data from Experiment 3, were sufficient to address the questions we had hoped to answer from these experiments.

@@

Dipper functions from Experiment 1 are displayed in the upper row of Figure \ref{fig:dipperfig}. Panel (a) shows the achromatic results, which replicate the key features from previous work. At detection threshold, binocular summation was a factor of 1.67 (4.47dB), within the range (\(\sqrt{2}\) to 2) consistent with previous reports (Baker et al., 2018). Pedestal masking functions followed the typical `dipper' shape in all conditions, with a region of facilitation at low pedestal contrasts, and masking at higher contrasts. The monocular and binocular dipper handles converge at high contrasts, whereas the half-binocular thresholds remain above the binocular thresholds across the full range of pedestal contrasts. The dichoptic condition produced very high thresholds, with the rising portion of the dipper having a slope around 1 (regression slope of 1.06 in log (dB) units, calculated across the highest 4 pedestal contrasts).

@@

{\centering \includegraphics{Figures/dipperssimplex}

@@

\caption{Dipper functions and psychometric slopes from Experiment 1, averaged across three participants. Panels (a-c) show threshold data, and panels (d-f) represent the slope of the psychometric function expressed in Weibull $\beta$ units. Error bars give ±1SE across participants. Note that contrast values are expressed as a percentage of the maximum displayable contrast (see Procedures for details). Curves in panels (a-c) show the best fitting models, optimized using a simplex algorithm (described in section 4.4), and RMSE values give the root mean square errors of the fits.}\label{fig:dipperfig}
\end{figure}

A similar pattern of results was observed for both the red/green and blue/yellow isoluminant stimuli (see Figure \ref{fig:dipperfig}b,c). Summation at threshold was a factor of 1.71 (4.66dB) for the red/green targets, and a factor of 1.73 (4.77dB) for the blue/yellow targets, and so was marginally higher than for achromatic stimuli. The general character of the dipper functions was largely consistent with the achromatic results, though we observed shallower facilitation and weaker masking, especially for the blue/yellow stimuli. For example, the strongest facilitation in the binocular condition for achromatic stimuli was a factor of 2.63, whereas it reduced to a factor of 2.35 for red/green stimuli and 1.57 for blue/yellow stimuli. The slope of the binocular dipper handle was 0.52 for achromatic stimuli, 0.58 for red/green stimuli, and 0.34 for blue/yellow stimuli. Dichoptic masking remained as strong for the chromatic conditions as for the achromatic stimuli (regression slopes of 1.3 for red/green and 1.21 for blue/yellow). The pattern of results for individual participants was consistent with the group averages, as shown in Figure \ref{fig:individualdips}.

Following Meese et al. (2006), we also inspected the slope of the psychometric function for each condition (see Figure \ref{fig:dipperfig}d-f). At detection threshold, slopes were relatively steep for all stimuli, with \(\beta\) values around 4. As pedestal contrasts increased, slopes linearized and reduced to around \(\beta=1.3\) (Foley \& Legge, 1981; Meese et al., 2006), and remained shallow at high pedestal contrasts. The exception to this was the dichoptic condition, where slopes became extremely steep at high dichoptic mask contrasts, consistent with previous observations (Baker et al., 2013; Meese et al., 2006). This was clear for all three data sets, with slope values in the range \(4 < \beta < 8\). However, we observe that slope estimates are more variable than threshold estimates (note the large error bars), particularly when using adaptive staircases, which deploy the majority of trials close to threshold. Our second experiment therefore investigated the slope of the psychometric function in more detail for dichoptic masking using the method of constant stimuli.

@@

In Experiment 2 we focussed on the dichoptic condition at a single mask contrast, and measured full psychometric functions using the method of constant stimuli for all factorial pairings of target and mask chromaticity. The pooled results across three participants are shown in Figure \ref{fig:MCSfig}a-c, and results for individual participants are available in Figure \ref{fig:individualMCS}. All conditions produced monotonically increasing psychometric functions (panels a-c), but the extent of masking was highly dependent on the relationship between the target and mask chromaticity. Figure \ref{fig:MCSfig}d-f shows a two-dimensional representation of individual threshold and slope estimates (points), as well as the posterior density estimates for fits to the pooled data (ellipses). The results are consistent between participants, and at the group level, and show that the presence of a mask has a strong effect on thresholds.

@@

{\centering \includegraphics{Figures/MCSdata}

@@

{\centering \includegraphics[width=0.8\linewidth]{Figures/discdata}

@@

\caption{Thresholds for the flickering disc experiment. Plotting conventions mirror those in Figure 4.}\label{fig:discdata}
\end{figure}

@@

{\centering \includegraphics{Figures/stanoutput}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  pdfauthor={Daniel H. Baker; Kirralise J. Hansford; Federico G. Segala; Annie Y. Morsi; Rowan J. Huxley; Joel T. Martin; Maya Rockman; Alex R. Wade},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Binocular integration of chromatic and luminance signals}
\author{Daniel H. Baker \and Kirralise J. Hansford \and Federico G.
Segala \and Annie Y. Morsi \and Rowan J. Huxley \and Joel T.
Martin \and Maya Rockman \and Alex R. Wade}
\date{}

\begin{document}
\maketitle
\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

@@

A useful framework for understanding binocular signal processing is the
two-stage gain control model of binocular combination introduced by
Meese et al. (2006). This model features interocular suppression between
monocular channels, followed by binocular summation. The model accounts
well for the pattern of contrast discrimination (`dipper') functions for
four distinct ocular configurations (see also Georgeson et al., 2016),
illustrated in Figure~\ref{fig-exampledips}. In the monocular condition,
participants must discriminate between stimuli of two contrasts (a
`pedestal', and a `pedestal plus target') that are both presented to one
eye, whilst the other eye views mean luminance. Threshold is defined as
the minimum target contrast required to make this judgement with 75\%
accuracy. The binocular condition is the same, except that the same
stimuli are shown to both eyes. In the half-binocular condition the
pedestal is shown to both eyes, but the target increment shown only to
one eye. Finally, the dichoptic condition involves presenting the
pedestal to one eye, and the target increment to the other eye.

@@

{\centering \includegraphics{Figures/exampledips.pdf}

@@

\caption{\label{fig-exampledips}Illustration of stimulus conditions
(left) and example dipper functions (right).}

\end{figure}

@@

In Experiments 1 and 2, the stimuli were horizontal sinusoidal gratings
with a spatial frequency of 1c/deg (see examples in
Figure~\ref{fig-examplestim}). The gratings were windowed by a raised
cosine envelope with a diameter of 3 degrees. Spatial phase, relative to
a central fixation cross, was randomised on each trial across the four
cardinal phases. In the achromatic conditions, the sine-wave modulated
all three monitor colour channels equally (red, green and blue). In the
L-M condition, we generated isoluminant stimuli for each participant
(see Procedures) designed to maximise contrast between L and M cones,
whilst keeping S cone activity constant. In the S-(L+M) condition, the
isoluminant stimuli maximised S-cone contrast. Stimuli were converted
from cone space to monitor RGB coordinates using the monitor spectral
readings and the Stockman-Sharpe 2 degree cone fundamentals (Stockman \&
Sharpe, 2000). The stimuli in Experiment 3 were temporal modulations of
luminance applied to a disc made using the same raised cosine envelope
as described above, but with no further spatial modulation. The stimuli
counterphase flickered sinusoidally at 4Hz (see examples in the lower
portion of Figure~\ref{fig-examplestim}). In all experiments, we
displayed a binocular fusion lock, consisting of three concentric rings
of small square elements with random colour. A black central fixation
cross was also displayed throughout.

@@

{\centering \includegraphics[width=0.5\textwidth,height=\textheight]{Figures/examplestim.pdf}

@@

\caption{\label{fig-examplestim}Example stimuli. Upper row shows grating
stimuli used in Experiments 1 and 2. Lower row shows one cycle of
sinusoidal flicker applied to a uniform disc. Note that the rendering of
all stimuli will depend on the device used to display or print this
image, and so the chromatic stimuli are unlikely to appear isoluminant,
and there may be additional luminance nonlinearities that were not
present in the stimuli displayed during the experiments.}

\end{figure}

@@

Before beginning primary data collection, each participant in
Experiments 1 and 2 completed an isoluminance adjustment task. Grating
stimuli were presented that counterphase flickered at 5Hz, defined about
either the L-M or S-(L+M) plane in cone space. Participants used a
trackball to dynamically adjust the colour angle of the stimulus to
minimise the percept of flicker. Each participant completed ten such
trials for each colour plane, and the average angle across repetition
was taken as the isoluminant point, and used to generate stimuli for the
main experiment for that participant. Settings were very similar across
participants for the S-(L+M) direction, and somewhat more heterogeneous
for the L-M direction (see Figure~\ref{fig-isofig}).

@@

{\centering \includegraphics{Figures/isosettings.pdf}

@@

\caption{\label{fig-isofig}Isoluminance settings from all participants
in Experiments 1 and 2. Panel (a) shows red/green and panel (b) shows
blue/yellow settings that were subsequently used to generate stimuli in
the main experiments. Within each panel, solid lines show the mean
settings for each participant, and black curves show the range of
possible stimuli displayed during the adjustment task.}

\end{figure}

In Experiment 1, participants completed a two-interval-forced-choice
(2IFC) contrast discrimination task. Stimuli were presented for 200ms,
with an interstimulus interval of 400ms. Each interval was indicated by
an auditory beep, and participants made their responses using a
two-button trackball. Correct responses were followed by a high pitched
tone, and incorrect responses by a low pitched tone. Each block of the
experiment tested a single pedestal contrast level, and lasted around 12
minutes. On each trial the target contrast level was determined by a
3-down-1-up staircase procedure. There were 8 interleaved staircases in
total; four stimulus arrangements (see Figure~\ref{fig-exampledips})
combined factorially with two target eye assignments. Each pedestal
contrast was repeated 3 times by each participant, and the block order
was randomised. The experiment lasted around 4 hours per participant for
each chromatic condition, and took place over the course of several
weeks. In total, the experiment consisted of 67978 trials (pooled across
participants).

@@

We also implemented a Bayesian hierarchical version of the model using
the Stan language (Carpenter et al., 2017). This used a binomial
generator function to model the proportion correct data at each target
level, and was fit simultaneously to all participants for a given
experiment, but separately for each chromatic condition of Experiment 1,
and the flickering disc data from Experiment 3 (i.e.~four fits in total,
as for the simplex fitting). Prior distributions were Gaussian, with
means determined from published values (see first row of
Table~\ref{tbl-parametertable}). This modelling primarily focuses on
examining posterior parameter distributions, rather than a model
comparison approach. We generated over 1 million posterior samples for
the model, and retained 10\% of them for plotting.

@@

where \emph{AC} represents the achromatic contrast, \emph{RG} represents
the red/green contrast, \emph{BY} represents the blue/yellow contrast,
and \(\omega_A\), \(\omega_R\) and \(\omega_B\) are the accompanying
weights of interocular suppression. There is an equivalent expression
for the right eye, and for each of the two isoluminant chromatic
pathways. To simplify the model, and avoid free parameters that are
poorly constrained by the data, we fixed several parameters at the
average values from the fits from Experiment 1
(Table~\ref{tbl-parametertable}), such that \(p=7.82\), \(q=6.4\),
\(m=1.21\), \(S=1.03\) and \(k=0.23\). We additionally fixed the
within-mechanism weight of interocular suppression (\(\omega\)) at 1.
This left 9 free parameters: a \(Z\) parameter for each mechanism, and
six cross-mechanism weights of interocular suppression. These parameters
were again estimated within a Bayesian hierarchical framework, using the
trial-wise data from Experiment 2. Note that the model as specified does
not currently include monocular suppression between different pathways,
as we did not collect any data for these conditions. Previous work (Chen
et al., 2000) has measured such interactions, and they could in
principle be incorporated into the denominator of either stage 1 or
stage 2 in the model.

@@

All experimental code, raw data and analysis scripts are available at:
https://osf.io/3vdga/. The linked GitHub repository also contains a
fully reproducible version of the manuscript. Note that we deviated
slightly from the planned preregistration, in that we did not collect
data for chromatic flickering discs, or for the cross-pathway dichoptic
experiment using disc stimuli. This is because the grating data from
Experiments 1 and 2, and the achromatic disc data from Experiment 3,
were sufficient to address the questions we had hoped to answer from
these experiments.

@@

Dipper functions from Experiment 1 are displayed in the upper row of
Figure~\ref{fig-dipperfig}. Panel (a) shows the achromatic results,
which replicate the key features from previous work. At detection
threshold, binocular summation was a factor of 1.67 (4.47dB), within the
range (\(\sqrt{2}\) to 2) consistent with previous reports (Baker et
al., 2018). Pedestal masking functions followed the typical `dipper'
shape in all conditions, with a region of facilitation at low pedestal
contrasts, and masking at higher contrasts. The monocular and binocular
dipper handles converge at high contrasts, whereas the half-binocular
thresholds remain above the binocular thresholds across the full range
of pedestal contrasts. The dichoptic condition produced very high
thresholds, with the rising portion of the dipper having a slope around
1 (regression slope of 1.06 in log (dB) units, calculated across the
highest 4 pedestal contrasts).

@@

{\centering \includegraphics{Figures/dipperssimplex.pdf}

@@

\caption{\label{fig-dipperfig}Dipper functions and psychometric slopes
from Experiment 1, averaged across three participants. Panels (a-c) show
threshold data, and panels (d-f) represent the slope of the psychometric
function expressed in Weibull \(\beta\) units. Error bars give ±1SE
across participants. Note that contrast values are expressed as a
percentage of the maximum displayable contrast (see Procedures for
details). Curves in panels (a-c) show the best fitting models, optimized
using a simplex algorithm (described in section 4.4), and RMSE values
give the root mean square errors of the fits.}

\end{figure}

A similar pattern of results was observed for both the red/green and
blue/yellow isoluminant stimuli (see Figure~\ref{fig-dipperfig}b,c).
Summation at threshold was a factor of 1.71 (4.66dB) for the red/green
targets, and a factor of 1.73 (4.77dB) for the blue/yellow targets, and
so was marginally higher than for achromatic stimuli. The general
character of the dipper functions was largely consistent with the
achromatic results, though we observed shallower facilitation and weaker
masking, especially for the blue/yellow stimuli. For example, the
strongest facilitation in the binocular condition for achromatic stimuli
was a factor of 2.63, whereas it reduced to a factor of 2.35 for
red/green stimuli and 1.57 for blue/yellow stimuli. The slope of the
binocular dipper handle was 0.52 for achromatic stimuli, 0.58 for
red/green stimuli, and 0.34 for blue/yellow stimuli. Dichoptic masking
remained as strong for the chromatic conditions as for the achromatic
stimuli (regression slopes of 1.3 for red/green and 1.21 for
blue/yellow). The pattern of results for individual participants was
consistent with the group averages, as shown in
Figure~\ref{fig-individualdippers}.

Following Meese et al. (2006), we also inspected the slope of the
psychometric function for each condition (see
Figure~\ref{fig-dipperfig}d-f). At detection threshold, slopes were
relatively steep for all stimuli, with \(\beta\) values around 4. As
pedestal contrasts increased, slopes linearized and reduced to around
\(\beta=1.3\) (Foley \& Legge, 1981; Meese et al., 2006), and remained
shallow at high pedestal contrasts. The exception to this was the
dichoptic condition, where slopes became extremely steep at high
dichoptic mask contrasts, consistent with previous observations (Baker
et al., 2013; Meese et al., 2006). This was clear for all three data
sets, with slope values in the range \(4 < \beta < 8\). However, we
observe that slope estimates are more variable than threshold estimates
(note the large error bars), particularly when using adaptive
staircases, which deploy the majority of trials close to threshold. Our
second experiment therefore investigated the slope of the psychometric
function in more detail for dichoptic masking using the method of
constant stimuli.

@@

In Experiment 2 we focussed on the dichoptic condition at a single mask
contrast, and measured full psychometric functions using the method of
constant stimuli for all factorial pairings of target and mask
chromaticity. The pooled results across three participants are shown in
Figure~\ref{fig-MCSfig}a-c, and results for individual participants are
available in Figure~\ref{fig-individualMCS}. All conditions produced
monotonically increasing psychometric functions (panels a-c), but the
extent of masking was highly dependent on the relationship between the
target and mask chromaticity. Figure~\ref{fig-MCSfig}d-f shows a
two-dimensional representation of individual threshold and slope
estimates (points), as well as the posterior density estimates for fits
to the pooled data (ellipses). The results are consistent between
participants, and at the group level, and show that the presence of a
mask has a strong effect on thresholds.

@@

{\centering \includegraphics{Figures/MCSdata.pdf}

@@

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/discdata.pdf}

@@

\caption{\label{fig-discdata}Thresholds for the flickering disc
experiment. Plotting conventions mirror those in Figure 4.}

\end{figure}

@@

{\centering \includegraphics{Figures/stanoutput.pdf}
